Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-10-04 forecast/paper/paper.tex
Add companion refs

\dropcap{T}racking and forecasting indicators from public health reporting
streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
crucial for understanding disease spread, formulating public policy responses,
and anticipating future public health resource needs.  In a companion paper, we
describe our research group's (Delphi's) efforts in curating and maintaining a 
database of real-time indicators that track COVID-19 activity and other relevant
phenomena. The signals (a term we use synonomously with ``indicators'') in this
database are accessible through the COVIDcast API \cite{CovidcastAPI}, with
associated R \cite{CovidcastR} and Python \cite{CovidcastPy} packages for
convenient data fetching and processing tools. In the current paper, we aim to
quantify the utility provided by a core set of these indicators for two
fundamental prediction tasks: probabilistic forecasting of COVID-19 case 
rates and prediction of future COVID-19 case hotspots (defined by the event that
a relative increase in COVID-19 cases exceeds a certain threshold).

@@

In our companion paper, we analyze correlations between various indicators and   
COVID case rates. These correlations are natural summaries of the
contemporaneous association between an indicator and COVID cases, but they fall 
short of delivering a satisfactory answer to the question that motivates the
current article: is the information contained in an indicator demonstrably
useful for the prediction tasks we care about? For such a question, lagged
correlations (e.g., measuring the correlation between an indicator and COVID
case rates several days in the future) move us in the right direction, but still
do not move us all the way there. The question about \textit{utility for 
  prediction} is focused on a much higher standard than simply asking 
about correlations; to be useful in forecast or hotspot models, an indicator
must provide relevant information that is not otherwise contained in past values
of the case rate series itself. We will assess this in the most direct way
possible: by inspecting the difference in predictive performance of simple
autoregressive models trained with and without access to past values of a
particular indicator.

@@

For more details on the five indicators (including how these are precisely
computed from the underlying data streams) we refer to
\url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html},
which documents all of the signals in the COVIDcast API, and our companion paper
on the API and database. For CTIS in particular, we refer to our companion paper
on this survey. For the Google COVID-19 Search Trends data set, see
\cite{GoogleSymptoms}; see also \cite{Klopfen:2020, Vaira:2020} for a 
justification of the relevance of anosmia or ageusia to COVID-19 infection.

@@

In this paper, all models are fit with ``vintage'' training data. This means
that for a given prediction date, say, September 28, 2020, we train models 
using data that would have been available to us \textit{as of}
September 28 (imagine that we can ``rewind'' the clock to September 28 and query
the COVIDcast API to get the latest data it would have had available at that
point in time.)  This is possible because of the COVIDcast API's comprehensive
data versioning system (described in more detail in our companion paper).  We
also use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the 
process of training arbitrary prediction models over a sequence of prediction
dates, by constructing the proper sequence of vintage training data sets.

>>>>

\dropcap{T}racking and forecasting indicators from public health reporting
streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
crucial for understanding disease spread, formulating public policy responses,
and anticipating future public health resource needs.  A companion paper
\cite{Reinhart:2021} describes our research group's (Delphi's) efforts in
curating and maintaining a database of real-time indicators that track COVID-19
activity and other relevant phenomena. The signals (a term we use synonomously
with ``indicators'') in this database are accessible through the COVIDcast API
\cite{CovidcastAPI}, with associated R \cite{CovidcastR} and Python
\cite{CovidcastPy} packages for convenient data fetching and processing
tools. In the current paper, we aim to quantify the utility provided by a core
set of these indicators for two fundamental prediction tasks: probabilistic
forecasting of COVID-19 case rates and prediction of future COVID-19 case
hotspots (defined by the event that a relative increase in COVID-19 cases
exceeds a certain threshold).

@@

In the companion paper \cite{Reinhart:2021}, we analyze correlations between
various indicators and COVID case rates. These correlations are natural
summaries of the contemporaneous association between an indicator and COVID
cases, but they fall short of delivering a satisfactory answer to the question
that motivates the current article: is the information contained in an indicator
demonstrably useful for the prediction tasks we care about? For such a question,
lagged correlations (e.g., measuring the correlation between an indicator and
COVID case rates several days in the future) move us in the right direction, but
still do not move us all the way there. The question about \textit{utility for
  prediction} is focused on a much higher standard than simply asking about 
correlations; to be useful in forecast or hotspot models, an indicator must
provide relevant information that is not otherwise contained in past values of
the case rate series itself. We will assess this in the most direct way
possible: by inspecting the difference in predictive performance of simple
autoregressive models trained with and without access to past values of a
particular indicator.

@@

For more details on the five indicators (including how these are precisely
computed from the underlying data streams) we refer to
\url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html},
and the companion paper on the COVIDcast API and its signals
\cite{Reinhart:2021}. For CTIS in particular, we refer to the companion paper
\cite{Salomon:2021}. For the Google COVID-19 Search Trends data set, see
\cite{GoogleSymptoms}; see also \cite{Klopfen:2020, Vaira:2020} for a
justification of the relevance of anosmia or ageusia to COVID infection.

@@

In this paper, all models are fit with ``vintage'' training data. This means
that for a given prediction date, say, September 28, 2020, we train models 
using data that would have been available to us \textit{as of}
September 28 (imagine that we can ``rewind'' the clock to September 28 and query
the COVIDcast API to get the latest data it would have had available at that
point in time.)  This is possible because of the COVIDcast API's comprehensive
data versioning system (described in more detail in the companion paper
\cite{Reinhart:2021}).  We also use the \texttt{evalcast} R package
\cite{EvalcastR}, which streamlines the process of training arbitrary
prediction models over a sequence of prediction dates, by constructing the
proper sequence of vintage training data sets.
